### 2015-05-05 background.tex
updates in the lockset analysis part of the background section

Example~\ref{fig:data_race_example} presents multiple data races that \whoop automatically found in the generic\_nvram Linux char driver (see Section~\ref{evaluation}). The driver can invoke the \texttt{nvram\_llseek} entry point concurrently from two threads $T_1$ and $T_2$, while passing the same \texttt{file} object as a parameter. However, this can lead to multiple possible data races because $T_1$ and $T_2$ can read/write access the \texttt{f\_pos} field of the \texttt{file} object in arbitrary order resulting into nondeterministic behavior.

@@

\ADComment{What is missing from the explanation, and from the example you give, is the idea that with a dynamic approach, you run the program and a thread operates in a truly concurrent environment.  You thus only get to see what the thread did on the particular schedule that was followed.  I also think you should use a two-threaded example, instead of a one-threaded example, to show that although thread $T$ might consistently use mutex $M$ to protect variable $A$, this is no good if thread $U$ accesses variable $A$ without holding $M$.}  \ADComment{If you use multiple letters in math mode, you should use ``mathit'' to get them typeset properly.  E.g., you should write $\mathit{CLS}_T$, not $CLS_T$, and $\mathit{LS}_v$, not $LS_v$.  Can you please address this throughout the paper?  It can be cumbersome to write, making it worth thinking about adding macros.}

Lockset analysis for a concurrent program starts by creating a \emph{current} lockset, $CLS_T$, for each thread $T$ of the program, and a lockset, $LS_v$, for each shared variable $v$ used in the program.
Initially, $CLS_T$ is empty for every thread $T$, because the threads do not hold any locks on program start.
The lockset $LS_v$ for each variable $v$ is initialised to the set of all locks manipulated by the program,
because initially each access to $v$ has (vacuously) been protected by every lock.
The program is executed as usual (with threads scheduled according to the OS scheduler), except that
instrumentation is added to update locksets as follows:
%
\begin{enumerate}
\item After each \emph{lock} and \emph{unlock} operation by thread $T$, $CLS_T$ is updated to reflect the locks currently held by $T$.
\item When thread $T$ accesses shared variable $v$, any locks that are not common to $LS_v$ and $CLS_T$ are removed from $LS_v$.
If $LS_v$ becomes empty as a result, a warning is issued that the access to $v$ may be unprotected.
\end{enumerate}

In the example of Figure~\ref{fig:locksets}, we analyze only one thread $T$ and one shared variable $A$ and thus there are two locksets: the current lockset for $T$ ($CLS_T$) and the lockset for $A$ ($LS_A$).

Figure~\ref{fig:locksets} shows an example of applying lockset analysis on a thread of a concurrent program.
\ADComment{I folded some of the discussion of the example into the above more general description of locksets.  Assuming you do produce a two-threaded example, please put in a brief explanation of it.}

Lockset analysis avoids reasoning about arbitrary thread interleavings \ADComment{is this true of standard lockset analysis?  With Eraser, a single interleaving would indeed provide a lot of information about races, but it would still be necessary to explore lots of interleavings to get high coverage, since the program behaviour may be schedule-dependent}, and thus has the potential to scale well in realistic concurrent programs. The technique, though, suffers from imprecision (i.e. can report many false bugs), because a violation of the locking discipline does not always correspond to a real data race~\cite{savage1997eraser, pozniansky2003efficient, o2003hybrid, elmas2007goldilocks, flanagan2009fasttrack}. Furthermore, as a classic dynamic analyzer, Eraser's bug-finding ability is limited to the execution paths that the tool explores. \ADComment{Perhaps re-phrase this to say that some existing techniques (cite them) have explored the idea of static lock set analysis, based on dataflow analysis, and that we also present a static lockset analysis method, that involves generating verification conditions to be discharged to a theorem prover} To counter these limitations, we apply the idea of Eraser's lockset analysis in a static verification context.

>>>>

Example~\ref{fig:data_race_example} presents multiple data races that \whoop automatically found in the generic\_nvram Linux char driver (see Section~\ref{evaluation}). The driver can invoke the \texttt{nvram\_llseek} entry point concurrently from two threads $T_1$ and $T_2$, while passing the same \texttt{file} object as a parameter. However, this can lead to multiple possible data races, because $T_1$ and $T_2$ can read/write access the \texttt{f\_pos} field of the \texttt{file} object in arbitrary order resulting into nondeterministic behavior.

@@

\ADComment{What is missing from the explanation, and from the example you give, is the idea that with a dynamic approach, you run the program and a thread operates in a truly concurrent environment.  You thus only get to see what the thread did on the particular schedule that was followed.  I also think you should use a two-threaded example, instead of a one-threaded example, to show that although thread $T$ might consistently use mutex $M$ to protect variable $A$, this is no good if thread $U$ accesses variable $A$ without holding $M$.}

Lockset analysis for a concurrent program starts by creating a \emph{current} lockset, $\mathit{CLS}_T$, for each thread $T$ of the program, and a lockset, $\mathit{LS}_v$, for each shared variable $v$ used in the program. Initially, $\mathit{CLS}_T$ is empty for every thread $T$, because the threads do not hold any locks on program start. The lockset $\mathit{LS}_v$ for each variable $v$ is initialized to the set of all locks manipulated by the program, because initially each access to $v$ has (vacuously) been protected by every lock. The program is executed as usual (with threads scheduled according to the OS scheduler), except that instrumentation is added to update locksets as follows:
%
\begin{enumerate}
\item After each \emph{lock} and \emph{unlock} operation by $T$, $\mathit{CLS}_T$ is updated to reflect the locks currently held by $T$.
\item When $T$ accesses shared variable $v$, any locks that are not common to $\mathit{LS}_v$ and $\mathit{CLS}_T$ are removed from $\mathit{LS}_v$.
If $\mathit{LS}_v$ becomes empty as a result, a warning is issued that the access to $v$ may be unprotected.
\end{enumerate}

In the example of Figure~\ref{fig:locksets}, we analyze only one thread $T$ and one shared variable $A$ and thus there are two locksets: the current lockset for $T$ ($\mathit{CLS}_T$) and the lockset for $A$ ($\mathit{LS}_A$).

Figure~\ref{fig:locksets} shows an example of applying lockset analysis on a concurrent program consisting of two threads $T_1$ and $T_2$, both accessing the shared variable $A$. Thus, we have to analyze three locksets: the current locksets $\mathit{CLS}_1$ and $\mathit{CLS}_2$ for $T_1$ and $T_2$ respectively; and the lockset $\mathit{LS}_A$ for $A$.

\ADComment{I folded some of the discussion of the example into the above more general description of locksets.  Assuming you do produce a two-threaded example, please put in a brief explanation of it.}

In contrast to analyses based on happens-before, lockset analysis has the potential to scale well, because it does not need to reason about all possible thread interleavings to achieve good code coverage. The technique, though, suffers from imprecision (i.e. can report false bugs), because a violation of the locking discipline does not always correspond to a real data race~\cite{savage1997eraser, pozniansky2003efficient, o2003hybrid, elmas2007goldilocks, flanagan2009fasttrack}. Furthermore, code coverage in dynamic lockset analysis tools, such as Eraser, is limited by the number of explored execution paths. To counter these limitations, some techniques~\cite{pratikakis2006locksmith, voung2007relay} have explored the idea of applying lockset analysis in a static context, using dataflow analysis. In this paper, we also present a static lockset analysis method that involves generating verification conditions to be discharged to a theorem prover.
